---
title: "Introduction to lpt: Local Parallel Trends"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to lpt: Local Parallel Trends}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Overview

The `lpt` package implements partial identification for continuous
difference-in-differences (DiD) designs under local restrictions on
counterfactual trends.

Parallel trends in continuous DiD requires the selection function
$\mu_t(d) = E[\Delta Y_t(0) \mid D = d]$ to be constant in dose. `lpt`
replaces that equality with two logically independent bounds, each matched
to a target parameter:

- **Level restriction (Assumption L):** $|\mu_t(d) - \mu_t(0)| \leq M$,
  which bounds treatment-effect *levels*:
  $$IS_{ATT}(d, t; M) = [\Lambda_t(d) - (t+1)M,\ \Lambda_t(d) + (t+1)M]$$
- **Slope restriction (Assumption S):** $|\mu'_t(d)| \leq B$ on the
  interior of the treated dose support, which bounds *dose-response
  derivatives*:
  $$IS_{\partial ATT}(d, t; B) = [\lambda_t(d) - (t+1)B,\ \lambda_t(d) + (t+1)B]$$
- **Overall summary:** the per-period $ATT^o_t$ has identified set
  $\pm(t+1)M$ around the binary DiD estimand, and the time-aggregated
  $ATT^o$ has half-width $\frac{T+2}{2}M$ for contiguous horizons $0..T$.

The sets widen linearly in the horizon $t$ because each post-period
accumulates one more unit of potential counterfactual drift. Setting
$M = 0$ (resp. $B = 0$) recovers point identification under standard
parallel trends for the corresponding target.

Levels get their own parameter $M$ — not a dose-proportional bound derived
from $B$ — because the dose support has a hole between 0 and the smallest
positive dose: a within-support slope bound cannot control the deviation
from the untreated group without extrapolating across that gap.

## Example: French Social Housing (SRU Law)

The bundled `sru` dataset contains panel data on 991 French communes.
Periods are re-indexed so treatment onset is period 0 (pre: -7 to -1,
post: 0 to 5).

```{r data}
library(lpt)
data(sru)
str(sru)
```

### Standard Parallel Trends (M = 0, B = 0)

Setting both bounds to zero recovers standard DiD --- point identification:

```{r pt0}
fit0 <- lpt(sru, "commune", "year", "outcome", "dose",
            post_period = 5, pre_periods = -7:-1, M = 0, B = 0)
fit0
```

### Calibrated M and B from Pre-Periods

Under no anticipation, pre-period first differences identify the selection
function: $E[\Delta Y_s \mid D = d] = \mu_s(d)$ for $s < 0$. Under stable
selection, the package calibrates
$\hat{M} = \max_s \sup_d |\hat\mu_s(d) - \hat\mu_s(0)|$ and
$\hat{B} = \max_s \sup_d |\hat\mu'_s(d)|$:

```{r cal, message = FALSE}
fit <- lpt(sru, "commune", "year", "outcome", "dose",
           post_period = 0:5, pre_periods = -7:-1,
           M = "calibrate", B = "calibrate")
summary(fit)
```

Note the practical asymmetry: $\hat{M}$ requires only conditional means of
pre-period first differences, whereas $\hat{B}$ requires their derivatives
--- a harder, noisier estimation problem. The level route is the more
robust default when only ATTs are of interest.

### Visualizations

```{r plots, fig.width = 8, fig.height = 5, message = FALSE}
library(ggplot2)

# Event study: ATT^o across pre and post periods (IS via M)
plot(fit, type = "eventstudy")

# Dose-response slope identified set (via B)
plot(fit, type = "datt")

# ATT level identified set (via M; constant width across doses)
plot(fit, type = "att")

# Pre-period diagnostics: trend deviations (M) and selection slopes (B)
plot(fit, type = "pretrends")

# Sensitivity of ATT^o in M
plot(fit, type = "sensitivity")
```

### Sensitivity Analysis

Both bounds accept vectors, computing identified sets for each value:

```{r sensitivity, message = FALSE}
fit_sens <- lpt(sru, "commune", "year", "outcome", "dose",
                post_period = 5, pre_periods = -7:-1,
                M = seq(0, 0.02, by = 0.005),
                B = seq(0, 0.05, by = 0.01))
summary(fit_sens)
```
