---
title: "Introduction to lpt: Local Parallel Trends"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to lpt: Local Parallel Trends}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Overview

The `lpt` package implements partial identification for continuous
difference-in-differences (DiD) designs under Local Parallel Trends.

In standard continuous DiD, the observable dose-slope decomposes as:

$$\lambda(d) = \frac{\partial ATT(d|d)}{\partial d} + \mu'(d)$$

where $\lambda(d)$ is identified from data but the selection slope $\mu'(d)$
is not. Under the Local Parallel Trends assumption, $|\mu'(d)| \leq B$,
which yields identified sets whose width grows with the post-treatment
horizon $t$:

- **Derivative:** $IS_{\partial ATT}(d, t; B) = [\lambda_t(d) - (t+1)B, \lambda_t(d) + (t+1)B]$
- **Level:** $IS_{ATT}(d, t; B) = [\Lambda_t(d) - (t+1)Bd, \Lambda_t(d) + (t+1)Bd]$
- **Overall:** $IS_{ATT^o_t}(B) = [ATT^o_{bin,t} - (t+1)B\bar{D}_+, ATT^o_{bin,t} + (t+1)B\bar{D}_+]$

## Example: French Social Housing (SRU Law)

The bundled `sru` dataset contains panel data on 991 French communes.
Periods are re-indexed so treatment onset is period 0 (pre: -7 to -1,
post: 0 to 5).

```{r data}
library(lpt)
data(sru)
str(sru)
```

### Standard Parallel Trends (B = 0)

Setting $B = 0$ recovers standard DiD --- point identification:

```{r b0}
fit0 <- lpt(sru, "commune", "year", "outcome", "dose",
            post_period = 5, pre_periods = -7:-1, B = 0)
fit0
```

### Calibrated B from Pre-Periods

The sensitivity parameter $B$ can be calibrated from pre-treatment selection
slopes $\hat{\mu}'_s(d)$:

```{r bcal, message = FALSE}
fit <- lpt(sru, "commune", "year", "outcome", "dose",
           post_period = 0:5, pre_periods = -7:-1,
           B = "calibrate")
summary(fit)
```

### Visualizations

```{r plots, fig.width = 8, fig.height = 5, message = FALSE}
library(ggplot2)

# Dose-response slope identified set
plot(fit, type = "datt")

# ATT level identified set
plot(fit, type = "att")

# Pre-period selection slopes
plot(fit, type = "pretrends")

# Sensitivity analysis
plot(fit, type = "sensitivity")
```

### Sensitivity Analysis

Run over a grid of B values:

```{r sensitivity, message = FALSE}
B_grid <- seq(0, 0.5, by = 0.05)
fit_sens <- lpt(sru, "commune", "year", "outcome", "dose",
                post_period = 5, pre_periods = -7:-1,
                B = B_grid)
summary(fit_sens)
```
